---
title: 'QMSS Data Viz - Assignment #2'
output: html_document
---

Julia Tache

Professor Thomas Brambor 

GR5063: QMSS Data Visualization 

03/30/20 

Assignment #2: Mapping Fires (and hopefully putting them out) 
------------------------------------------------------------------

```{r}
library(tidyverse) 
library(ggmap)

# fire_all <- read_csv("Incidents_Responded_to_by_Fire_Companies.csv")
# fire <- fire_all %>%
#   filter(INCIDENT_TYPE_DESC == "111 - Building fire")

# Register Google API Key
# register_google(key = "AIzaSyB-D2Q8JlnM8XET9G8D4roPEpEhHd6ubhE", write = TRUE)

# Create and geocode addresses
# fire <- fire %>%
#   mutate(address = str_c( str_to_title(fire$STREET_HIGHWAY),
#                   "New York, NY",
#                   fire$ZIP_CODE,
#                   sep=", ")) %>%
#   filter(is.na(address)==FALSE) %>%
#   mutate_geocode(address)

# Save File
# write_csv(fire, "building_fires.csv")
```

#### 1. Location of Severe Fires

```{r}
library(tidyverse)
library(leaflet) 
library(sp)
library(maps) 

firehouses <- read_csv("FDNY_Firehouse_Listing.csv") %>%
  dplyr::filter(!is.na(Latitude))
fire <- read.csv("building_fires.txt") 
fire <-rename(fire, "lng" = "lon")
fire <- fire %>% filter(BOROUGH_DESC == c("1 - Manhattan", "2 - Bronx", "3 - Staten Island", 
                                          "4 - Brooklyn", "5 - Queens"))

severe_fire <- fire %>% filter(HIGHEST_LEVEL_DESC == 	c("7 - Signal 7-5"))
# subsets by the highest level fire 

bins <- c(0, 5, 10, 15, 20, 30, Inf)
pal <- colorBin(c("light yellow", "yellow", "orange", "red", "dark red", "black"), 
                domain = severe_fire$UNITS_ONSCENE, bins = bins)

map <- leaflet(data = severe_fire) %>% 
  addTiles() %>%
  addCircles(lng = ~lng, 
             lat = ~lat, 
             color = ~pal(UNITS_ONSCENE), 
             stroke =TRUE, 
             fillOpacity = 0.5,
             radius = ~(as.numeric(FIRE_SPREAD_DESC)*50),
             popup = paste("Incident Type:", severe_fire$INCIDENT_TYPE_DESC, "<br>",
                           "Date and Time:",severe_fire$INCIDENT_DATE_TIME, "<br>",
                           "Fire Spread:", severe_fire$FIRE_SPREAD_DESC)) %>%
  addLegend("bottomright", pal = pal, values = ~UNITS_ONSCENE,
            title = "Number of Units on Scene",
            opacity = 1)

map

```

#### 2. Layers and Clusters

##### a) Color by Type of Property

```{r}

# clean the data and collapse property types into main categories 
severe_fire$PROPERTY_USE_DESC <- str_replace_all(severe_fire$PROPERTY_USE_DESC, 
                                                 "24-hour care Nursing homes, 4 or more persons", 
                                                 "Twenty-Four Hour Nursing Homes") 

severe_fire$PROPERTY_USE_DESC <- str_replace_all(severe_fire$PROPERTY_USE_DESC, 
                                                 "1 or 2", "One or two") 

severe_fire$PROPERTY_USE_DESC <- str_replace_all(severe_fire$PROPERTY_USE_DESC, 
                                                 "UUU - Undetermined", "000") 

severe_fire$PROPERTY_USE_DESC <- str_replace_all(severe_fire$PROPERTY_USE_DESC, 
                                                 "\\D", "") 

severe_fire$PROPERTY_USE_DESC <- gsub("000", "Other", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("1[0-9][0-9]", "Recreation and Religion", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("2[0-9][0-9]", "Daycare", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("3[0-9][0-9]", "Health and Safety", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("4[0-9][0-9]", "Housing", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("5[0-9][0-9]", "Business", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("6[0-9][0-9]", "Laboratory", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("7[0-9][0-9]", "Manufacturing", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("8[0-9][0-9]", "Storage/Fire Stations", severe_fire$PROPERTY_USE_DESC) 
severe_fire$PROPERTY_USE_DESC <- gsub("9[0-9][0-9]", "Open Land/Lots", severe_fire$PROPERTY_USE_DESC) 

pal <- colorFactor(c("Paired"), domain = severe_fire$PROPERTY_USE_DESC)

map <- leaflet(data = severe_fire) %>% 
  addTiles() %>%
  addCircles(lng = ~lng, 
             lat = ~lat, 
             color = ~pal(PROPERTY_USE_DESC), 
             stroke =TRUE, 
             fillOpacity = 0.5,
             radius = ~(as.numeric(FIRE_SPREAD_DESC)*50),
             popup = paste("Property Type:", severe_fire$PROPERTY_USE_DESC, "<br>",
                           "Date and Time:",severe_fire$INCIDENT_DATE_TIME, "<br>",
                           "Fire Spread:", severe_fire$FIRE_SPREAD_DESC)) %>%
  addLegend("bottomright", pal = pal, values = ~PROPERTY_USE_DESC,
            title = "Property Use",
            opacity = 1)

map
```

##### b) Cluster

```{r}
map <- map %>% addMarkers(clusterOptions = markerClusterOptions())

map
```

#### 3. Fire Houses

```{r}
firehouse_icon <- makeIcon(
  iconUrl = "https://www.pngrepo.com/png/45189/180/fire-station.png",
  iconWidth = 25, iconHeight = 25,
  iconAnchorX = 22, iconAnchorY = 94,
)

map <- leaflet(data = severe_fire) %>% 
  addTiles(group = "OSM (default)") %>%
  addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
  addProviderTiles(providers$Stamen.TonerLite, group = "Toner Lite") %>%
  addCircles(lng = ~lng, 
             lat = ~lat, 
             color = ~pal(PROPERTY_USE_DESC), 
             stroke = TRUE, 
             fillOpacity = 0.5,
             radius = ~(as.numeric(UNITS_ONSCENE)*10),
             group = "Incidents",
             popup = paste("Property Type:", severe_fire$PROPERTY_USE_DESC, "<br>",
                           "Date and Time:",severe_fire$INCIDENT_DATE_TIME, "<br>",
                           "UNITS_ONSCENE:", severe_fire$UNITS_ONSCENE)) %>%
  addLegend("bottomright", pal = pal, values = ~PROPERTY_USE_DESC,
            title = "Property Use",
            opacity = 1) %>%
  addMarkers(data = firehouses, 
              group = "Firehouses",     
              icon = firehouse_icon, 
              popup = paste("Facility Name:", firehouses$FacilityName)) %>%
  addLayersControl(
    baseGroups = c("OSM (default)", "Toner", "Toner Lite"),
    overlayGroups = c("Incidents", "Firehouses"),
    options = layersControlOptions(collapsed = FALSE)
  )

map

```

Firehouse Icon Source: https://www.pngrepo.com/png/45189/180/fire-station.png 

#### 4. Distance from Firehouse and Response Time

##### a) Calculate Distance

```{r}
library(geosphere)
library(lubridate)

# calculate shortest distance between fire and firehouse  
fire <- mutate(fire, 
       `Distance in Meters` = distHaversine(cbind(fire$lng, fire$lat),
                                cbind(firehouses$Longitude, firehouses$Latitude), r = 6378137))

fire$ARRIVAL_DATE_TIME <- mdy_hms(fire$ARRIVAL_DATE_TIME) 
fire$INCIDENT_DATE_TIME <- mdy_hms(fire$INCIDENT_DATE_TIME) 

# calculate response time 
fire <- mutate(fire, `Response Time in Minutes` = 
                 difftime(fire$ARRIVAL_DATE_TIME, fire$INCIDENT_DATE_TIME) / 60)

fire$`Response Time in Minutes` <- as.numeric(fire$`Response Time in Minutes`) 
fire$`Response Time in Minutes` <- round(fire$`Response Time in Minutes`, 3)

library(ggplot2)

ggplot(fire, aes(x =`Response Time in Minutes`, y = `Distance in Meters`)) + 
  geom_point() + xlim(0, 20) + ylim(0, 50000)

fire3 <- filter(fire, HIGHEST_LEVEL_DESC == c("0 - Initial alarm", 
                                              "1 - More than initial alarm, less than Signal 7-5", 
                                              "7 - Signal 7-5")) 

ggplot(fire3, aes(x =`Response Time in Minutes`, y = `Distance in Meters`, 
                  color = HIGHEST_LEVEL_DESC)) + 
  geom_point() + xlim(0, 7.5) + ylim(0, 50000) 

```

I had hypothesized that there would be a clear difference between the response time for high severity fires and low severity fires, but this does not seem to be the case. The scatterplot shows that there are similar response times for both Alarm 1 and Alarm 7 fires, with a slight upward trend of response times and the distance of the fire to the nearest firehouse. The response times overall, excluding outliers which are not shown in the graphs above, were very quick (2-10 minutes). A lot can happen in 2-10 minutes of a fire, so hopefully some of these less severe fires are able to be put out quickly and not become higher level fires. 

##### b) Map of Response Times

```{r}
library(tigris)
library(dplyr)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)

# open data map of nyc neighborhoods 
nyc_neighborhoods <- readOGR("nyc_neighborhoods.json")

# binds existing fire data frame with nyc neighborhoods 
fire_spdf <- fire
coordinates(fire_spdf) <- ~lng + lat
proj4string(fire_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(fire_spdf, nyc_neighborhoods)
points <- cbind(fire, matches)

# summarizing average response time by neighborhood for easy mapping 
time_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(Average = round(mean(`Response Time in Minutes`, na.rm = TRUE), digits = 3)) 

map_data <- geo_join(nyc_neighborhoods, time_by_neighborhood, "neighborhood", "neighborhood")

bins <- c(0, 1, 2, 3, 4, 5, 10, Inf) 
pal <- colorBin("RdYlBu", domain = map_data$Average, bins = bins)

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(Average), 
              popup = paste("Neighborhood:", map_data$neighborhood, "<br>",
                            "Average Response Time (in minutes):", map_data$Average), 
              opacity = 1,
              color = "black",
              dashArray = "1", 
              fillOpacity = 1,
              weight = 1, 
              highlight = highlightOptions(
                weight = 5,
                color = "white",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
                addLegend(pal = pal, values = ~Average, opacity = 0.7,
                          position = "bottomright", title = "Average Response Time in Minutes") %>%
                addMarkers(data = fire, 
                           lng = ~lng, 
                           lat = ~lat, 
                           group = "Incidents",
                           popup = paste("Property Type:", fire$PROPERTY_USE_DESC, "<br>",
                                         "Response Time (in minutes):", fire$`Response Time in Minutes`, "<br>",
                                         "Level of Fire:", fire$HIGHEST_LEVEL_DESC)) %>% 
  addLayersControl(
                  baseGroups = c("OSM (default)"),
                  overlayGroups = c("Incidents"),
                  options = layersControlOptions(collapsed = FALSE))
```

Based on the markers, it does not seem like the property type or level of fire really play much of a role in the response time. However, the map showcases how distance does play a role, as we can see neighborhoods out in Queens and Staten Island have sometimes the longest wait times of 8+ minutes. 

```{r}
fire$INCIDENT_DATE_TIME <- str_sub(fire$INCIDENT_DATE_TIME, 1, 4)

fire_13 <- filter(fire, INCIDENT_DATE_TIME == "2013")
fire_14 <- filter(fire, INCIDENT_DATE_TIME == "2014")
fire_15 <- filter(fire, INCIDENT_DATE_TIME == "2015")
fire_16 <- filter(fire, INCIDENT_DATE_TIME == "2016")
fire_17 <- filter(fire, INCIDENT_DATE_TIME == "2017")
fire_18 <- filter(fire, INCIDENT_DATE_TIME == "2018")

nyc_neighborhoods <- readOGR("nyc_neighborhoods.json")

fire_spdf_13 <- fire_13
coordinates(fire_spdf_13) <- ~lng + lat
proj4string(fire_spdf_13) <- proj4string(nyc_neighborhoods)
matches <- over(fire_spdf_13, nyc_neighborhoods)
points <- cbind(fire_13, matches)

time_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(Average = round(mean(`Response Time in Minutes`, na.rm = TRUE), digits = 3)) 

map_data <- geo_join(nyc_neighborhoods, time_by_neighborhood, "neighborhood", "neighborhood")

bins <- c(0, 1, 2, 3, 4, 5, 10, Inf) 
pal <- colorBin("RdYlBu", domain = map_data$Average, bins = bins)

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(Average), 
              popup = paste("Neighborhood:", map_data$neighborhood, "<br>",
                            "Average Response Time (in minutes):", map_data$Average), 
              opacity = 1,
              color = "black",
              weight = 1, 
              dashArray = "1",
              fillOpacity = 1,
              highlight = highlightOptions(
                weight = 5,
                color = "white",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
                addLegend(pal = pal, values = ~Average, opacity = 0.7,
                          position = "bottomright", title = "Average Response Time in Minutes (2013)") %>%
                addMarkers(data = fire_13, 
                           lng = ~lng, 
                           lat = ~lat, 
                           group = "Incidents",
                           popup = paste("Property Type:", fire_13$PROPERTY_USE_DESC, "<br>",
                                         "Response Time (in minutes):", fire_13$`Response Time in Minutes`, "<br>",
                                         "Level of Fire:", fire_13$HIGHEST_LEVEL_DESC)) %>% 
  addLayersControl(
                  baseGroups = c("OSM (default)"),
                  overlayGroups = c("Incidents"),
                  options = layersControlOptions(collapsed = FALSE))

```

```{r}
nyc_neighborhoods <- readOGR("nyc_neighborhoods.json")

fire_spdf_14 <- fire_14
coordinates(fire_spdf_14) <- ~lng + lat
proj4string(fire_spdf_14) <- proj4string(nyc_neighborhoods)
matches <- over(fire_spdf_14, nyc_neighborhoods)
points <- cbind(fire_14, matches)

time_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(Average = round(mean(`Response Time in Minutes`, na.rm = TRUE), digits = 3)) 

map_data <- geo_join(nyc_neighborhoods, time_by_neighborhood, "neighborhood", "neighborhood")

bins <- c(0, 1, 2, 3, 4, 5, 10, Inf) 
pal <- colorBin("RdYlBu", domain = map_data$Average, bins = bins)

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(Average), 
              popup = paste("Neighborhood:", map_data$neighborhood, "<br>",
                            "Average Response Time (in minutes):", map_data$Average), 
              opacity = 1,
              color = "black",
              dashArray = "1",
              weight = 1, 
              fillOpacity = 1,
              highlight = highlightOptions(
                weight = 5,
                color = "white",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
                addLegend(pal = pal, values = ~Average, opacity = 0.7,
                          position = "bottomright", title = "Average Response Time in Minutes (2014)") %>%
                addMarkers(data = fire_14, 
                           lng = ~lng, 
                           lat = ~lat, 
                           group = "Incidents",
                           popup = paste("Property Type:", fire_14$PROPERTY_USE_DESC, "<br>",
                                         "Response Time (in minutes):", fire_14$`Response Time in Minutes`, "<br>",
                                         "Level of Fire:", fire_14$HIGHEST_LEVEL_DESC)) %>% 
  addLayersControl(
                  baseGroups = c("OSM (default)"),
                  overlayGroups = c("Incidents"),
                  options = layersControlOptions(collapsed = FALSE))
```

```{r}
nyc_neighborhoods <- readOGR("nyc_neighborhoods.json")

fire_spdf_15 <- fire_15
coordinates(fire_spdf_15) <- ~lng + lat
proj4string(fire_spdf_15) <- proj4string(nyc_neighborhoods)
matches <- over(fire_spdf_15, nyc_neighborhoods)
points <- cbind(fire_15, matches)

time_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(Average = round(mean(`Response Time in Minutes`, na.rm = TRUE), digits = 3)) 

map_data <- geo_join(nyc_neighborhoods, time_by_neighborhood, "neighborhood", "neighborhood")

bins <- c(0, 1, 2, 3, 4, 5, 10, Inf) 
pal <- colorBin("RdYlBu", domain = map_data$Average, bins = bins)

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(Average), 
              popup = paste("Neighborhood:", map_data$neighborhood, "<br>",
                            "Average Response Time (in minutes):", map_data$Average), 
              opacity = 1,
              color = "black",
              dashArray = "1",
              weight = 1, 
              fillOpacity = 1,
              highlight = highlightOptions(
                weight = 5,
                color = "white",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
                addLegend(pal = pal, values = ~Average, opacity = 0.7,
                          position = "bottomright", title = "Average Response Time in Minutes (2015)") %>%
                addMarkers(data = fire_15, 
                           lng = ~lng, 
                           lat = ~lat, 
                           group = "Incidents",
                           popup = paste("Property Type:", fire_15$PROPERTY_USE_DESC, "<br>",
                                         "Response Time (in minutes):", fire_15$`Response Time in Minutes`, "<br>",
                                         "Level of Fire:", fire_15$HIGHEST_LEVEL_DESC)) %>% 
  addLayersControl(
                  baseGroups = c("OSM (default)"),
                  overlayGroups = c("Incidents"),
                  options = layersControlOptions(collapsed = FALSE))
```

```{r}
nyc_neighborhoods <- readOGR("nyc_neighborhoods.json")

fire_spdf_16 <- fire_16
coordinates(fire_spdf_16) <- ~lng + lat
proj4string(fire_spdf_16) <- proj4string(nyc_neighborhoods)
matches <- over(fire_spdf_16, nyc_neighborhoods)
points <- cbind(fire_16, matches)

time_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(Average = round(mean(`Response Time in Minutes`, na.rm = TRUE), digits = 3)) 

map_data <- geo_join(nyc_neighborhoods, time_by_neighborhood, "neighborhood", "neighborhood")

bins <- c(0, 1, 2, 3, 4, 5, 10, Inf) 
pal <- colorBin("RdYlBu", domain = map_data$Average, bins = bins)

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(Average), 
              popup = paste("Neighborhood:", map_data$neighborhood, "<br>",
                            "Average Response Time (in minutes):", map_data$Average), 
              opacity = 1,
              color = "black",
              dashArray = "1",
              weight = 1, 
              fillOpacity = 1,
              highlight = highlightOptions(
                weight = 5,
                color = "white",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
                addLegend(pal = pal, values = ~Average, opacity = 0.7,
                          position = "bottomright", title = "Average Response Time in Minutes (2016)") %>%
                addMarkers(data = fire_16, 
                           lng = ~lng, 
                           lat = ~lat, 
                           group = "Incidents",
                           popup = paste("Property Type:", fire_16$PROPERTY_USE_DESC, "<br>",
                                         "Response Time (in minutes):", fire_16$`Response Time in Minutes`, "<br>",
                                         "Level of Fire:", fire_16$HIGHEST_LEVEL_DESC)) %>% 
  addLayersControl(
                  baseGroups = c("OSM (default)"),
                  overlayGroups = c("Incidents"),
                  options = layersControlOptions(collapsed = FALSE))
```

```{r}
nyc_neighborhoods <- readOGR("nyc_neighborhoods.json")

fire_spdf_17 <- fire_17
coordinates(fire_spdf_17) <- ~lng + lat
proj4string(fire_spdf_17) <- proj4string(nyc_neighborhoods)
matches <- over(fire_spdf_17, nyc_neighborhoods)
points <- cbind(fire_17, matches)

time_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(Average = round(mean(`Response Time in Minutes`, na.rm = TRUE), digits = 3)) 

map_data <- geo_join(nyc_neighborhoods, time_by_neighborhood, "neighborhood", "neighborhood")

bins <- c(0, 1, 2, 3, 4, 5, 10, Inf) 
pal <- colorBin("RdYlBu", domain = map_data$Average, bins = bins)

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(Average), 
              popup = paste("Neighborhood:", map_data$neighborhood, "<br>",
                            "Average Response Time (in minutes):", map_data$Average), 
              opacity = 1,
              color = "black",
              dashArray = "1",
              fillOpacity = 1,
              weight = 1, 
              highlight = highlightOptions(
                weight = 5,
                color = "white",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
                addLegend(pal = pal, values = ~Average, opacity = 0.7,
                          position = "bottomright", title = "Average Response Time in Minutes (2017)") %>%
                addMarkers(data = fire_17, 
                           lng = ~lng, 
                           lat = ~lat, 
                           group = "Incidents",
                           popup = paste("Property Type:", fire_17$PROPERTY_USE_DESC, "<br>",
                                         "Response Time (in minutes):", fire_17$`Response Time in Minutes`, "<br>",
                                         "Level of Fire:", fire_17$HIGHEST_LEVEL_DESC)) %>% 
  addLayersControl(
                  baseGroups = c("OSM (default)"),
                  overlayGroups = c("Incidents"),
                  options = layersControlOptions(collapsed = FALSE))

```

```{r}
nyc_neighborhoods <- readOGR("nyc_neighborhoods.json")

fire_spdf_18 <- fire_18
coordinates(fire_spdf_18) <- ~lng + lat
proj4string(fire_spdf_18) <- proj4string(nyc_neighborhoods)
matches <- over(fire_spdf_18, nyc_neighborhoods)
points <- cbind(fire_18, matches)

time_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(Average = round(mean(`Response Time in Minutes`, na.rm = TRUE), digits = 3)) 

map_data <- geo_join(nyc_neighborhoods, time_by_neighborhood, "neighborhood", "neighborhood")

bins <- c(0, 1, 2, 3, 4, 5, 10, Inf) 
pal <- colorBin("RdYlBu", domain = map_data$Average, bins = bins)

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(Average), 
              popup = paste("Neighborhood:", map_data$neighborhood, "<br>",
                            "Average Response Time (in minutes):", map_data$Average), 
              opacity = 1,
              color = "black",
              dashArray = "1",
              fillOpacity = 1,
              weight = 1, 
              highlight = highlightOptions(
                weight = 5,
                color = "white",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
                addLegend(pal = pal, values = ~Average, opacity = 0.7,
                          position = "bottomright", title = "Average Response Time in Minutes (2018)") %>%
                addMarkers(data = fire_18, 
                           lng = ~lng, 
                           lat = ~lat, 
                           group = "Incidents",
                           popup = paste("Property Type:", fire_18$PROPERTY_USE_DESC, "<br>",
                                         "Response Time (in minutes):", fire_18$`Response Time in Minutes`, "<br>",
                                         "Level of Fire:", fire_18$HIGHEST_LEVEL_DESC)) %>% 
  addLayersControl(
                  baseGroups = c("OSM (default)"),
                  overlayGroups = c("Incidents"),
                  options = layersControlOptions(collapsed = FALSE))
```

Over time (2013 - 2018), response times seem to have gotten faster given the increased amount of warm colors on the map, but there seems to have been a slowdown of response times from 2017-2018. There are also a lot of neighborhoods with "NAs" especially in later years in the data, which I am hoping means that there were less fires that the FDNY responded to as the years went on. Some trends remain consistent: heavily populated areas/areas close to the city center (Manhattan) seem to have faster response times, while fires out on the edges of the boroughs tend to have units arrive on the scene later. 

Citations: 

https://www.r-graph-gallery.com/327-chloropleth-map-from-geojson-with-ggplot2.html

https://rpubs.com/jhofman/nycmaps 

https://rstudio.github.io/leaflet/

